Reliable error estimation for Sobol' indices\thanks{This work is supported by the CITiES project funded by the Agence Nationale de la Recherche (grant ANR-12-MONU-0020) and by the United States National Science Foundation (grant DMS-1522687).}

### 2017-02-16 article_review/Manuscript.tex
changes manuscript

\subsection{Estimation of Sobol' indices}
\label{sec:2.2}
In this section we review two Monte Carlo procedures for the estimation of Sobol' indices. A design is a point set $\mathcal{P}=\{\vx_i\}_{i=0}^{n-1}$ where each point is obtained by sampling each variable $x_j$ $n$ times. Each row of the design is a point $\vx_i$ in $[0,1]^d$ and each column of the design refers to samples of a variable $x_j$. Consider $\mathcal{P}=\{\vx_i\}_{i=0}^{n-1}$ and $\mathcal{P'}=\{{\vx'}_i\}_{i=0}^{n-1}$ two designs where $(\vx_i,{\vx'}_i) \iid \mathcal{U}([0,1]^{2d})$. One way to estimate the two quantities (\ref{first.order}) and (\ref{total.effect}) is via:
\begin{align}
\label{first.order.est}
\widehat{\underline{\tau}_u^2} &= \frac{1}{n} \sum \limits_{i=0}^{n-1} \left(f(\vx_i)-f(\vx_{i,u}:{\vx'}_{i,-u})\right)f(\vx'_i),\\
\label{total.effect.est}
\widehat{\ov{$\tau$}_u^2} &= \frac{1}{2n} \sum \limits_{i=0}^{n-1} (f({\vx'}_i) - f(\vx_{i,u}:{\vx'}_{i,-u}))^2, \qquad u \in \mathcal{D},
\end{align}
using for $\sigma^2$:
\begin{equation}
 \widehat{\sigma}^2 = \frac{1}{n-1} \sum \limits_{i=0}^{n-1} \left( f(\vx_i) - \widehat{\mu}\right)^2, \text{ with } \ \widehat{\mu} =  \frac{1}{n} \sum \limits_{i=0}^{n-1} f(\vx_i).
\label{mu.est}
\end{equation}
Then, the Sobol' indices estimators are:
\begin{equation}
\widehat{\underline{S}}_u = \widehat{\underline{\tau}_u^2} / \widehat{\sigma}^2, \qquad
\widehat{\overline{S}}_u = \widehat{\overline{\tau}_u^2} / \widehat{\sigma}^2.
\label{common.sobol.est}
\end{equation}
Using this method, the estimation of a single pair ($\widehat{\underline{S}}_u$, $\widehat{\overline{S}}_u$) requires $3n$ evaluations of the model $f$. Thus, for all first order and total effect indices, one would need $3nd$ model evaluations. Using a combinatorial formalism, in \cite{Saltelli} Saltelli proposes the following estimation strategy:
\begin{theorem}
\label{saltelli.theorem}
The $d+2$ designs $\{\vx_{i,u}:{\vx'}_{i,-u}\}_{i=0}^{n-1}$ constructed for $u \in \{\varnothing,\{1\},\dots,$ $\{d\},\mathcal{D}\}$ allow to estimate all first-order and total effect Sobol' indices at a cost of $n(d+2)$ evaluations of the model.
\end{theorem}
The main idea of the theorem is that there is no need to reevaluate $f({\vx}_i)$ and $f({\vx'}_i)$ for each $u$. Hence, we can simply evaluate $f({\vx}_i)$ $n$ times, $f({\vx'}_i)$ $n$ times, and $f(\vx_{i,u}:{\vx'}_{i,-u})$ $nd$ times, which accounts for the $n(d+2)$ evaluations.

@@

A more efficient alternative to evaluate all first-order indices was proposed in \cite{Mara} and only requires $2n$ model evaluations. This alternative relies on the construction of two replicated designs. The notion of replicated designs was first introduced by McKay through his replicated Latin Hypercubes in \cite{McKay}. In order to apply this definition to other types of points, in \cite{GJAHMP} we generalized this notion to the following.
\begin{definition}
\label{rep.designs}
Let $\mathcal{P}=\{\vx_i\}_{i=0}^{n-1}$ and $\mathcal{P}'=\{{\vx'}_i\}_{i=0}^{n-1}$ be two point sets in
$[0,1]^{d}$. Let $\mathcal{P}^u=\{\vx_{i,u}\}_{i=0}^{n-1}$ (resp. ${\mathcal{P}'}^u$), $u \subsetneq \mathcal{D}$, denote the subset of dimensions of $\mathcal{P}$ (resp. $\mathcal{P}'$) indexed by $u$. We say that $\mathcal{P}$ and $\mathcal{P}'$ are two replicated designs of order $a \in \{1,\dots,d-1\}$ if for any $u \subsetneq \mathcal{D}$ such that $|u|=a$, $\mathcal{P}^u$ and ${\mathcal{P}'}^u$ are the same point set in $[0,1]^a$. We define by $\pi_u$ the permutation that rearranges the rows of ${\mathcal{P}'}^u$ into $\mathcal{P}^u$.
\end{definition}
The method introduced in \cite{Mara} allows to estimate all first-order Sobol' indices with only two replicated designs of order $1$. The key point of this method is to use the permutations resulting from the structure of the two replicated designs to mimic the hybrid points in equation (\ref{first.order.est}).

@@

Finally, we define the set of functions $\cc$,
\begin{multline} \label{conecond}
\cc:=\{f \in \mathbb{L}^2(\cube) : \hS_{\ell,m}(f) \le \homega(m-\ell) \wcS_m(f),\ \ \ell \le m, \\
\wcS_m(f) \le \wcomega(m-\ell) S_{\ell}(f),\ \  \ell_* \le \ell \le m\}.
\end{multline}
for a fixed $\ell_* \in \mathbb{N}$, $\homega$ and $\wcomega$ two non-negative valued functions with $\lim_{m \to \infty} \wcomega(m) = 0$. \comment{The decay rate assumptions on the Walsh coefficients are encoded in $\cc$ and controlled by $\homega$ and $\wcomega$. First, we assume that $\wcS_m(f)$ bounds $\hS_{\ell,m}(f)$ in some sense, in particular it bounds $\hS_{0,m}(f)$. This means that we can use $\wcS_m(f)$ as a good indicator for the error bound. Then, we also need $S_{\ell}(f)$ to bound $\wcS_m(f)$ which helps transforming the infinite sum into a finite sum. In this step, we are assuming that if $S_{\ell}(f)$ is small enough, we expect the high frequency coefficients, i.e. $\wcS_m(f)$, to also be small. By defining $r=m-\ell$, and assuming that $\homega(r)\wcomega(r)<1$ with $r\in\mathbb{N}$, one may see $r$ as a distance between $S_{\ell}(f)$ and $\wcS_m(f)$: larger values of $r$ will imply a smaller error bound but more dependence on smaller Walsh coefficients. Finally, we can write the error bound in terms of $\tS_{\ell,m}(f)$ instead of $\wcS_m(f)$ by using \eqref{discrite_true_relationship}.} In \cite[Sec. 4.2]{HicJim} it is shown that for any $f\in\cc$,
\begin{equation}
\absolute{I - \frac{1}{b^m}\sum_{\vx\in\mathcal{P}_m}f(\vx)}
\le \tS_{\ell,m}(f)\underbrace{\frac{\homega(m) \wcomega(r)}{1 - \homega(r) \wcomega(r)}}_{\mathfrak{C}(m)}=\epsilon_{\widehat{I}}, \label{errbd}
\end{equation}
where one may increase $m$ until the error bound $\epsilon_{\widehat{I}}$ is small enough.

@@

Algorithm \ref{recursive.algorithm} summarizes the main steps of our sequential procedure. First, one must fix the tolerance $\epsilon >0$ at which Sobol' indices must be estimated, and set $m=\ell_*+r$, where $\ell_*$ and $r$ are the two hyper-parameters defined in \eqref{conecond}. Then, one must construct the two designs $\mathcal{P}_{m}=\{\vx_i\}_{i=0}^{2^{m}-1}$ and $\mathcal{P}'_{m}=\{\vx'_i\}_{i=0}^{2^{m}-1}$ at Step $5$ according to the multiplicative approach detailed in \cite{GJAHMP}. The choice of $\ell_*$ must be large enough to ensure that each integrand of $\underline{S}_u$ and $\overline{S}_u$ (Section \ref{sec:3.2}) is in $\cc$ (see  (\ref{conecond})).

Under this construction, $\mathcal{P}_{m}$ and $\mathcal{P}'_{m}$ correspond to the first $2^{m}$ points of two independent Sobol' sequences. Furthermore, they possess the structure of two replicated designs of order $1$, as shown in \cite{GJAHMP}. In addition, $\mathcal{P}_{m}$ and $\mathcal{P}'_{m}$ can also be scrambled as long as both point sets share the same scrambling dimensionwise.
\bigskip

@@

Alternatively, $\mathcal{P}_{m}$ and $\mathcal{P}'_{m}$ can be used with the replication procedure to estimate all first-order Sobol' indices. This option is referred as Variant $B$ in Algorithm \ref{recursive.algorithm}.

@@

Assume that the number of small first-order indices is known and equals $\gamma$. We denote by $u_1,\dots,u_{\gamma}$ the indices of the corresponding inputs and $\Gamma = \{1,\dots,\gamma\}$. The cost of Variant $A$ including ``Correlation 2" becomes, 
\begin{equation}
\sum \limits_{j \in \Gamma} 2^{m''_{u_j}} + \sum \limits_{j \in \Gamma} 2^{m'_{u_j}} + \sum \limits_{j \in \mathcal{D}\setminus{\Gamma}} 2^{m_{u_j}} + 2 \times 2^{m^\star},
\label{cost.improvement}
\end{equation}
where: %$m'_{u_j}\leq m^{\star}$, $\ m_{u_j} \leq m^{\star}$, and: 
\begin{itemize}
\item[$\bullet$] for $j\in \Gamma$, $2^{m''_{u_j}}$ is the number of evaluations \\ $f\left(\vz_{i,u_j}:\vx_{i,-u_j}\right)$ to estimate $\underline{S}_{u_j}$,
\item[$\bullet$] for $j\in \Gamma$, $2^{m'_{u_j}}$ is the number of evaluations \\ $f\left(\vx_{i,u_j}:{\vx'}_{i,-u_j}\right)$ to estimate both $\underline{S}_{u_j}$ and  $\overline{S}_{u_j}$,
\item[$\bullet$] likewise, for $j\in \mathcal{D}\setminus{\Gamma}$, $ 2^{m_{u_j}}$ is the number of evaluations $f\left(\vx_{i,u_j}:{\vx'}_{i,-u_j}\right)$ to estimate both $\underline{S}_{u_j}$ and  $\overline{S}_{u_j}$,
\item[$\bullet$] $2 \times 2^{m^{\star}}$ is the number of evaluations $f\left(\vx_{i}\right)$ and $f\left({\vx'}_{i}\right)$ used in the estimation of each first-order and total effect index.
\end{itemize}
We recall the cost of Variant $A$ without  ``Correlation 2",
\begin{equation}
\sum \limits_{j \in \Gamma} 2^{m_{u_j}} + \sum \limits_{j \in \mathcal{D}\setminus{\Gamma}} 2^{m_{u_j}} + 2 \times 2^{m^{\star}}. 
%\qquad m_{u_j} \leq m^{\star}.
\label{cost.non.improvement}
\end{equation}
The difference (\ref{cost.improvement}) $-$ (\ref{cost.non.improvement}) equals:
\begin{equation}
 \sum \limits_{j \in \Gamma} 2^{m_{u_j}} \left(2^{m''_{u_j}-m_{u_j}}  + 2^{m'_{u_j}-m_{u_j}} - 1 \right) = \sum \limits_{j \in \Gamma} c_j.
\label{cost.comparison}
\end{equation}
Hence, the sign of this difference indicates whether or not using ``Correlation 2" brings an improvement to Variant $A$. We distinguish two cases :
\begin{itemize}
\item[1)] for $j \in \Gamma$, if the total effect index $\ov{$S$}_{u_j}$ requires as much or more evaluations than the first-order index $\underline{S}_{u_j}$. Since the total effect estimator is the same, as a consequence we have $m'_{u_j}=m_{u_j}$ and $c_j >0$.
\item[2)] for $j \in \Gamma$, if the total effect index $\ov{$S$}_{u_j}$ requires less evaluations than the first-order index $\underline{S}_{u_j}$. In this case, if both $m''_{u_j} < m_{u_j}$ and $m'_{u_j} < m_{u_j}$ then $c_j \leq 0$.  
\end{itemize}
Overall we expect to observe case $2)$ more often than case $1)$. Indeed, the numerator of $\underline{S}_u$ requires to estimate $2d$ dimensional integrals against only $d+1$ dimensional integrals for the numerator of $\ov{$S$}_u$. Hence, it seems reasonable to expect that it will take less points to estimate $\ov{$S$}_u$ than $\underline{S}_u$.

Furthermore, in case $2)$, we expect the two conditions $m''_{u_j} < m_{u_j}$ and $m'_{u_j} < m_{u_j}$ to usually hold as in \cite{Owen}, ``Correlation 2" is shown to perform better for small first-order indices. To support this discussion, numerical examples are presented in Section \ref{sec:5}.

In practice, one does not know which are the small Sobol' indices ($u_1,\dots,u_\gamma)$. To overcome this issue, we propose the following alternative for Variant $A$. If at the end of the first iteration $\widehat{\underline{S}}_u$ is smaller than a threshold (in our case $0.1$, as suggested by Owen), then the estimator (\ref{first.order.est}) is switched to ``Correlation 2" for this particular $u$ and a third Sobol' sequence $\mathcal{P''}_{m}=\{\vz_i\}_{i=0}^{2^m-1}$ is constructed to obtain the corresponding evaluations $f({\vz}_{i,u}:{\vx}_{i,-u})$.

@@

For the two test functions, results of the three variants are compared based on the true estimation errors. These errors correspond to the absolute difference between the true values and their estimates:
$$\delta_{\underline{S}_u} = \absolute{\underline{S}_u - \widehat{\underline{S}_u}}, \qquad \delta_{\overline{S}_u} = \absolute{\overline{S}_u - \widehat{\overline{S}_u}}.$$
We also compare the total number of evaluations of each variant. Results are averaged over $100$ repetitions using Owen's scrambled Sobol' sequences \cite{HonHic00a,Owe95}. For each repetition, we fix the tolerance $\varepsilon=5.10^{-3}$, set $\ell_*=5$, and $r=4$ (the algorithm starts with $512$ Sobol' points). The algorithm error bound factor described in \eqref{errbd} is set to $\mathfrak{C}(m)=10\times 2^{-m}$.

@@

@@

The two classical test functions considered in this article are the g-function introduced by Sobol' \cite{Sobol'}, and the function introduced by Bratley \textit{et al.} \cite{Bratley}. The idea is to test our method over two categories of functions: additive (Bratley \textit{et al.} function) and multiplicative (g-function).

@@

The g-function is defined as follows:
\begin{equation*}
f(\vx)= \prod \limits_{j=1}^{d}g_j(x_j), \qquad g_j(x_j)=\frac{\left |4x_j-2 \right | + a_j}{1 + a_j}, \ a_j \geq 0.
\end{equation*}
Each value $a_j$ determines the relative importance of the $x_j$. When the value of $a_j$ gets closer to zero the variable $x_j$ becomes more influent. For this example, we chose $d=6$ and $a_1=0,a_2=0.5, a_3=3, a_4=9, a_5=99$, and $a_6=99$.

>>>>

\subsection{Estimation of Sobol' indices}
\label{sec:2.2}
In this section we review two Monte Carlo procedures for the estimation of Sobol' indices. A design is a point set $\mathcal{P}=\{\vx_i\}_{i=0}^{n-1}$ where each point is obtained by sampling each variable $x_j$ $n$ times. Each row of the design is a point $\vx_i$ in $[0,1]^d$ and each column of the design refers to samples of a variable $x_j$. Consider $\mathcal{P}=\{\vx_i\}_{i=0}^{n-1}$ and $\mathcal{P'}=\{{\vx'}_i\}_{i=0}^{n-1}$ two designs where $(\vx_i,{\vx'}_i) \iid \mathcal{U}([0,1]^{2d})$. One way to estimate the two quantities (\ref{first.order}) and (\ref{total.effect}) is via:
\begin{align}
\label{first.order.est}
\widehat{\underline{\tau}_u^2} &= \frac{1}{n} \sum \limits_{i=0}^{n-1} \left(f(\vx_i)-f(\vx_{i,u}:{\vx'}_{i,-u})\right)f(\vx'_i),\\
\label{total.effect.est}
\widehat{\ov{$\tau$}_u^2} &= \frac{1}{2n} \sum \limits_{i=0}^{n-1} (f({\vx'}_i) - f(\vx_{i,u}:{\vx'}_{i,-u}))^2, \qquad u \in \mathcal{D},
\end{align}
using for $\sigma^2$:
{\color{purple} (Lolo)
\begin{equation}
 \widehat{\sigma}^2 = \frac{1}{n-1} \sum \limits_{i=0}^{n-1} \left( f(\vx_i) - \widehat{\mu}\right)^2, \text{ with } \ \widehat{\mu} =  \frac{1}{n} \sum \limits_{i=0}^{n-1} f(\vx_i).
\label{mu.est}
\end{equation}}
Then, the Sobol' indices estimators are:
\begin{equation}
\widehat{\underline{S}}_u = \widehat{\underline{\tau}_u^2} / \widehat{\sigma}^2, \qquad
\widehat{\overline{S}}_u = \widehat{\overline{\tau}_u^2} / \widehat{\sigma}^2.
\label{common.sobol.est}
\end{equation}
Using this method, the estimation of a single pair ($\widehat{\underline{S}}_u$, $\widehat{\overline{S}}_u$) requires $3n$ evaluations of the model $f$. Thus, for all first order and total effect indices, one would need $3nd$ model evaluations. Using a combinatorial formalism, in \cite{Saltelli} Saltelli proposes the following estimation strategy:
\begin{theorem}
\label{saltelli.theorem}
The $d+2$ designs $\{\vx_{i,u}:{\vx'}_{i,-u}\}_{i=0}^{n-1}$ constructed for $u \in \{\varnothing,\{1\},\dots,$ $\{d\},\mathcal{D}\}$ allow to estimate all first-order and total effect Sobol' indices at a cost of $n(d+2)$ evaluations of the model.
\end{theorem}
The main idea of the theorem is that there is no need to reevaluate $f({\vx}_i)$ and $f({\vx'}_i)$ for each $u$. Hence, we can simply evaluate $f({\vx}_i)$ $n$ times, $f({\vx'}_i)$ $n$ times, and $f(\vx_{i,u}:{\vx'}_{i,-u})$ $nd$ times, which accounts for the $n(d+2)$ evaluations.

@@

A more efficient alternative to evaluate all first-order indices was proposed in \cite{Mara} and only requires $2n$ model evaluations. This alternative relies on the construction of two replicated designs. The notion of replicated designs was first introduced by McKay through his replicated Latin Hypercubes in \cite{McKay}. In order to apply this definition to other types of points, in \cite{GJAHMP} we generalized this notion to the following.

{\color{purple} (Lolo) 
\begin{definition}
\label{rep.designs}
Let $\mathcal{P}=\{\vx_i\}_{i=0}^{n-1}$ and $\mathcal{P}'=\{{\vx'}_i\}_{i=0}^{n-1}$ be two point sets in $[0,1]^{d}$. Assume that the points $\vx_i$ (resp. $\vx'_i$) are all coordinate-wise different; that is $\forall (i_1,i_2) \in \{0,,n-1\}^2$, such that $i_1 \neq i_2$: $$x_{i_1,j} \neq x_{i_2,j}, \qquad \forall j=1,\dots,d .$$
Let $\mathcal{P}^u=\{\vx_{i,u}\}_{i=0}^{n-1}$ (resp. ${\mathcal{P}'}^u$), $u \subsetneq \mathcal{D}$, denote the subset of dimensions of $\mathcal{P}$ (resp. $\mathcal{P}'$) indexed by $u$. We say that $\mathcal{P}$ and $\mathcal{P}'$ are two replicated designs of order $a \in \{1,\dots,d-1\}$ if for any $u \subsetneq \mathcal{D}$ such that $|u|=a$, $\mathcal{P}^u$ and ${\mathcal{P}'}^u$ are the same point set in $[0,1]^a$. We define by $\pi_u$ the permutation that rearranges the rows of ${\mathcal{P}'}^u$ into $\mathcal{P}^u$.
\end{definition}

\begin{remark}
In the following, the two replicated designs $\mathcal{P}$, $\mathcal{P}'$ are constructed to avoid the degenerate case $\forall i \ \vx_i = \vx'_i$. 
\end{remark}}

The method introduced in \cite{Mara} allows to estimate all first-order Sobol' indices with only two replicated designs of order $1$. The key point of this method is to use the permutations resulting from the structure of the two replicated designs to mimic the hybrid points in equation (\ref{first.order.est}).

@@

Finally, we define the set of functions $\cc$,
\begin{multline} \label{conecond}
\cc:=\{f \in \mathbb{L}^2(\cube) : \hS_{\ell,m}(f) \le \homega(m-\ell) \wcS_m(f),\ \ \ell \le m, \\
\wcS_m(f) \le \wcomega(m-\ell) S_{\ell}(f),\ \  \ell_* \le \ell \le m\}.
\end{multline}
for a fixed $\ell_* \in \mathbb{N}$, $\homega$ and $\wcomega$ two non-negative valued functions with $\lim_{m \to \infty} \wcomega(m) = 0$. \comment{The decay rate assumptions on the Walsh coefficients are encoded in $\cc$ and controlled by $\homega$ and $\wcomega$. First, we assume that $\wcS_m(f)$ bounds $\hS_{\ell,m}(f)$ in some sense, in particular it bounds $\hS_{0,m}(f)$. This means that we can use $\wcS_m(f)$ as a good indicator for the error bound. Then, we also need $S_{\ell}(f)$ to bound $\wcS_m(f)$ which helps transforming the infinite sum into a finite sum. In this step, we are assuming that if $S_{\ell}(f)$ is small enough, we expect the high frequency coefficients, i.e. $\wcS_m(f)$, to also be small. By defining $r=m-\ell$, and assuming that $\homega(r)\wcomega(r)<1$ with $r\in\mathbb{N}$, one may see $r$ as a distance between $S_{\ell}(f)$ and $\wcS_m(f)$: larger values of $r$ will imply a smaller error bound but more dependence on smaller Walsh coefficients. Finally, we can write the error bound in terms of $\tS_{\ell,m}(f)$ instead of $\wcS_m(f)$ by using \eqref{discrite_true_relationship}.} In \cite[Sec. 4.2]{HicJim} it is shown that for any $f\in\cc$,
\begin{equation}
\absolute{I - \frac{1}{b^m}\sum_{\vx\in\mathcal{P}_m}f(\vx)}
\le \tS_{\ell,m}(f)\underbrace{\frac{\homega(m) \wcomega(r)}{1 - \homega(r) \wcomega(r)}}_{\mathfrak{C}(m)}=\epsilon_{\widehat{I}}, \label{errbd}
\end{equation}
where one may increase $m$ until the error bound {\color{purple}$\epsilon_{\widehat{I}}$} is small enough.

@@

Algorithm \ref{recursive.algorithm} summarizes the main steps of our sequential procedure. First, one must fix the tolerance $\epsilon >0$ at which Sobol' indices must be estimated, and set $m=\ell_*+r$, where $\ell_*$ and $r$ are the two hyper-parameters defined in \eqref{conecond}.  {\color{purple} (Lolo) The choice of $\ell_*$ must be large enough to ensure that each integrand of $\underline{S}_u$ and $\overline{S}_u$ (Section \ref{sec:3.2}) is in $\cc$ (see  (\ref{conecond})).}

Then, one must construct the two designs $\mathcal{P}_{m}=\{\vx_i\}_{i=0}^{2^{m}-1}$ and $\mathcal{P}'_{m}=\{\vx'_i\}_{i=0}^{2^{m}-1}$ at Step $5$ according to the multiplicative approach detailed in \cite{GJAHMP}. {\color{purple} (Lolo) This approach iteratively constructs a $2d$-dimensional Sobol' sequence then assigns its first $d$ coordinates to $\mathcal{P}_{m}$ and its last $d$ coordinates to $\mathcal{P}'_{m}$. The $2d$-dimension Sobol' sequence is generated based on the directional numbers found in \cite{Kuo}. These directional numbers insure better two-dimensional projections of the resulting Sobol' sequence. In addition, $\mathcal{P}_{m}$ and $\mathcal{P}'_{m}$ are also independently scrambled using Owen's scrambling scheme \cite{HonHic00a,Owe95}.}

%Under this construction, $\mathcal{P}_{m}$ and $\mathcal{P}'_{m}$ correspond to the first $2^{m}$ points of two independent Sobol' sequences. Furthermore, they possess the structure of two replicated designs of order $1$, as shown in \cite{GJAHMP}. In addition, $\mathcal{P}_{m}$ and $\mathcal{P}'_{m}$ can also be scrambled as long as both point sets share the same scrambling dimensionwise.
\bigskip

@@

Alternatively, $\mathcal{P}_{m}$ and $\mathcal{P}'_{m}$ can be used with the replication procedure to estimate all first-order Sobol' indices. This option is referred as Variant $B$ in Algorithm \ref{recursive.algorithm}. {\color{purple} (Lolo) However, in this case, the scrambling applied to both designs must be identical to insure that $\mathcal{P}_{m}$ and $\mathcal{P}'_{m}$ are replicated designs of order one.}

@@

Assume that the number of small first-order indices is known and equals $\gamma$. We denote by $u_1,\dots,u_{\gamma}$ the indices of the corresponding inputs and $\Gamma = \{1,\dots,\gamma\}$. The cost of Variant $A$ including ``Correlation 2" becomes, 
\begin{equation}
\sum \limits_{j \in \Gamma} 2^{m''_{u_j}} + \sum \limits_{j \in \Gamma} 2^{m'_{u_j}} + \sum \limits_{j \in \mathcal{D}\setminus{\Gamma}} 2^{m_{u_j}} + 2 \times 2^{m^\star},
\label{cost.improvement}
\end{equation}
where: %$m'_{u_j}\leq m^{\star}$, $\ m_{u_j} \leq m^{\star}$, and: 
\begin{itemize}
\item[$\bullet$] for $j\in \Gamma$, $2^{m''_{u_j}}$ is the number of evaluations \\ $f\left(\vz_{i,u_j}:\vx_{i,-u_j}\right)$ to estimate $\underline{S}_{u_j}$,
\item[$\bullet$] for $j\in \Gamma$, $2^{m'_{u_j}}$ is the number of evaluations \\ $f\left(\vx_{i,u_j}:{\vx'}_{i,-u_j}\right)$ to estimate both $\underline{S}_{u_j}$ and  $\overline{S}_{u_j}$,
\item[$\bullet$] likewise, for $j\in \mathcal{D}\setminus{\Gamma}$, $ 2^{m_{u_j}}$ is the number of evaluations $f\left(\vx_{i,u_j}:{\vx'}_{i,-u_j}\right)$ to estimate both $\underline{S}_{u_j}$ and  $\overline{S}_{u_j}$,
\item[$\bullet$] $2 \times 2^{m^{\star}}$ is the number of evaluations $f\left(\vx_{i}\right)$ and $f\left({\vx'}_{i}\right)$ used in the estimation of each first-order and total effect index.
\end{itemize}
We recall the cost of Variant $A$ without  ``Correlation 2",
\begin{equation}
\sum \limits_{j \in \Gamma} 2^{m_{u_j}} + \sum \limits_{j \in \mathcal{D}\setminus{\Gamma}} 2^{m_{u_j}} + 2 \times 2^{m^{\star}}. 
%\qquad m_{u_j} \leq m^{\star}.
\label{cost.non.improvement}
\end{equation}
The difference (\ref{cost.improvement}) $-$ (\ref{cost.non.improvement}) equals:
\begin{equation}
 \sum \limits_{j \in \Gamma} 2^{m_{u_j}} \left(2^{m''_{u_j}-m_{u_j}}  + 2^{m'_{u_j}-m_{u_j}} - 1 \right) = \sum \limits_{j \in \Gamma} c_j.
\label{cost.comparison}
\end{equation}
{\color{purple} (Lolo) The sign of \eqref{cost.comparison}} indicates whether or not using ``Correlation 2" brings an improvement to Variant $A$. We distinguish two cases :
\begin{itemize}
\item[1)] for $j \in \Gamma$, if the total effect index $\ov{$S$}_{u_j}$ requires as much or more evaluations than the first-order index $\underline{S}_{u_j}$. Since the total effect estimator is the same, as a consequence we have $m'_{u_j}=m_{u_j}$ and $c_j >0$.
\item[2)] for $j \in \Gamma$, if the total effect index $\ov{$S$}_{u_j}$ requires less evaluations than the first-order index $\underline{S}_{u_j}$. In this case, if both $m''_{u_j} < m_{u_j}$ and $m'_{u_j} < m_{u_j}$ then $c_j \leq 0$.  
\end{itemize}
{\color{purple}(Lolo) We assume that the effective dimension of the numerator of $\underline{S}_u$ is higher than the one of the numerator of $\ov{$S$}_u$. If this assumption holds, we expect to observe case $2)$ more often than case $1)$ and the sign of \eqref{cost.comparison} should be negative or equal to $0$. The validity of this assumption is tested on the numerical examples presented in Section \ref{sec:5}.
%Overall we expect to observe case $2)$ more often than case $1)$. Indeed, the numerator of $\underline{S}_u$ requires to estimate $2d$ dimensional integrals against only $d+1$ dimensional integrals for the numerator of $\ov{$S$}_u$. Hence, it seems reasonable to expect that it will take less points to estimate $\ov{$S$}_u$ than $\underline{S}_u$. 
Furthermore, in case $2)$, we expect the two conditions $m''_{u_j} < m_{u_j}$ and $m'_{u_j} < m_{u_j}$ to usually hold as ``Correlation 2" is shown to perform better for small first-order indices (\textit c.f. \cite{Owen}).}

In practice, one does not know which are the small Sobol' indices ($u_1,\dots,u_\gamma)$. To overcome this issue, we propose the following alternative for Variant $A$. {\color{purple}(Lolo) If at the end of the first iteration $\widehat{\underline{S}}_u$ is smaller than a threshold (in our case $0.1$, as suggested by Owen), the estimator (\ref{first.order.est}) is switched to ``Correlation 2" for this particular $u$. Then, a third Sobol' sequence $\mathcal{P''}_{m}=\{\vz_i\}_{i=0}^{2^m-1}$ is constructed to obtain the corresponding evaluations $f({\vz}_{i,u}:{\vx}_{i,-u})$. $\mathcal{P''}$ is obtained by assigning and scrambling the next $d$ coordinates of the starting $2d$-dimensional Sobol' sequence.}

@@

For the two test functions, results of the three variants are compared based on the true estimation errors. These errors correspond to the absolute difference between the true values and their estimates:
$$\delta_{\underline{S}_u} = \absolute{\underline{S}_u - \widehat{\underline{S}_u}}, \qquad \delta_{\overline{S}_u} = \absolute{\overline{S}_u - \widehat{\overline{S}_u}}.$$
{\color{purple} (Lolo) True values of Sobol' indices are estimated with a precision of $10$ digits. For clarity concern, they are rounded in the tables presenting the results.}

We also compare the total number of evaluations of each variant. {\color{purple} (Lolo) Results are averaged over $100$ repetitions by applying different scrambling}. For each repetition, we fix the tolerance $\varepsilon=5.10^{-3}$, set $\ell_*=5$, and $r=4$ (the algorithm starts with $512$ Sobol' points). The algorithm error bound factor described in \eqref{errbd} is set to $\mathfrak{C}(m)=10\times 2^{-m}$.

@@

{\color{purple} (Lolo) To verify the validity of our  assumption saying that the sign of \eqref{cost.comparison} is negative, the sum of the $c_j$ is evaluated for each repetition and barplots of its values are drawn.}

@@

The two classical test functions considered in this article are {\color{purple} (Lolo) a non QMC friendly version of the} g-function introduced by Sobol' \cite{Sobol'}, and the function introduced by Bratley \textit{et al.} \cite{Bratley}. The idea is to test our method over two categories of functions: additive (Bratley \textit{et al.} function) and multiplicative (g-function).

@@

{\color{purple} (Lolo)
The non QMC friendly version of the g-function is defined as follows:
\begin{equation*}
f(\vx)= \prod \limits_{j=1}^{d}g_j(x_j), \qquad g_j(x_j)=\frac{\left |3x_j-2 \right | + a_j}{1 + a_j}, \ a_j \geq 0.
\end{equation*}}
Each value $a_j$ determines the relative importance of the $x_j$. When the value of $a_j$ gets closer to zero the variable $x_j$ becomes more influent. For this example, we chose $d=6$ and $a_1=0,a_2=0.5, a_3=3, a_4=9, a_5=99$, and $a_6=99$.

Table \ref{res.gfunc.Aa} shows the averaged estimation errors obtained with Variant $A.a$ as well as the averaged total number of evaluations performed. Table \ref{res.gfunc.Ab} shows the same results obtained with Variant $A.b$.
\renewcommand{\arraystretch}{1.25}
{\color{purple}(Lolo) (nouvelles tables et figures)
\begin{table}[!ht]
\caption{Averaged estimation errors $\delta_{\underline{S}_u}$ , $\delta_{\overline{S}_u}$ and total number of evaluations for Variant $A.a$.}
\centering
\begin{tabular}{ccccccc}
\hline
 input & $\underline{S}_u$ & $\delta_{\underline{S}_u}$ & $r_{\delta_{\underline{S}_u}}$ & $\overline{S}_u$ & $\delta_{\overline{S}_u}$ &$r_{\delta_{\overline{S}_u}}$ \\ \hline
 $x_1$ & $0.6043$ &  $0.0005$ & $0$ & $0.7252$ & $0.0001$ & $0$ \\ \hline
 $x_2$ & $0.2360$ &  $0.0006$ & $0$ & $0.3481$ & $0.0007$ & $0$ \\ \hline
 $x_3$ & $0.0286$ &  $0.0007$ & $0$ & $0.0483$ & $0.0005$ & $0$ \\ \hline
 $x_4$ & $0.0043$ &  $0.0025$ & $0.11$ & $0.0075$ & $0.0004$ & $0$ \\ \hline
 $x_5$ & $4.10^{-5}$ & $0.0003$ & $0$ & $7.10^{-5}$ & $< 10^{-4}$ & $0$ \\ \hline
 $x_6$ & $4.10^{-5}$ & $0.0003$ & $0$  & $7.10^{-5}$ & $< 10^{-4}$ & $0$ \\ \hline \hline
\multicolumn{4}{l}{Total number of evaluations: $120 \ 638$} & & &\\ \hline 
\end{tabular}
\label{res.gfunc.Aa}
\end{table}
\begin{table}[!ht]
\caption{Averaged estimation errors $\delta_{\underline{S}_u}$ , $\delta_{\overline{S}_u}$ and total number of evaluations for Variant $A.b$.}
\centering
\begin{tabular}{ccccccc}
\hline
 input & $\underline{S}_u$ & $\delta_{\underline{S}_u}$ & $r_{\delta_{\underline{S}_u}}$ & $\overline{S}_u$ & $\delta_{\overline{S}_u}$ &$r_{\delta_{\overline{S}_u}}$ \\ \hline
 $x_1$ & $0.6043$ &  $0.0005$ & $0$ & $0.7252$ & $0.0001$ & $0$ \\ \hline
 $x_2$ & $0.2360$ &  $0.0006$ & $0$ & $0.3481$ & $0.0007$ & $0$ \\ \hline
 $x_3$ & $0.0286$ &  $0.0013$ & $0$ & $0.0483$ & $0.0005$ & $0$ \\ \hline
 $x_4$ & $0.0043$ &  $0.0003$ & $0$ & $0.0075$ & $0.0004$ & $0$ \\ \hline
 $x_5$ & $4.10^{-5}$ & $< 10^{-4}$ & $0$ & $7.10^{-5}$ & $< 10^{-4}$ & $0$ \\ \hline
 $x_6$ & $4.10^{-5}$ & $< 10^{-4}$ & $0$  & $7.10^{-5}$ & $< 10^{-4}$ & $0$ \\ \hline \hline
\multicolumn{4}{l}{Total number of evaluations: $120 \ 832$} & & &\\ \hline 
\end{tabular}
\label{res.gfunc.Ab}
\end{table}}
